Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-04-25 src/tex/ms.tex
fix figures again

% Abstract with filler text
\begin{abstract}
    We present a simple and general method of recalibrating gravitational wave
    (GW) phenomenological waveform models jointly. By using {\jax} and
    {\ripple}, we can perform automatic differentiation to functions, which
    allows us to use gradient-based optimization methods to recalibrate waveform
    coefficients in IMRPhenomD model. This method reduces systematic bias
    previously introduced to the model and generally can improve waveform
    accuracy. With recalibrated coefficients, we found that the typical
    \textit{mismatch} has a $50\%$ decrease. Furthermore, we analyze the
    accuracy base on the waveform's intrinsic parameters. We found that waveform
    accuracy has significant dependence on black hole spin. Reduced spin
    approximation introduces degeneracy in spin, which prevented further
    improvement. We isolated regions in the parameter space that does not fit
    the waveform ansatz. These results allow us to understand more about how to
    develop newer phenomenological models. 
\end{abstract}

% Main body with filler text
\section{Introduction} \label{sec:intro}

@@

\begin{figure}[t]
	\script{q148.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/q148.pdf}
	\caption{Distributions of waveform mismatches calculated using both
	$\mathcal{L}_{\mathrm{mean}}$ and $\mathcal{L}_{\mathrm{fl}}$. We use 
	training waveforms listed in Tab.~\ref{tab:q148} and
	mismatches are weighted with the constant PSD. For the
	$\mathcal{L}_{\mathrm{mean}}$ distributions, the median decreased by 50.0\%
	while the median of $\mathcal{L}_{\mathrm{fl}}$ distribution decreased by
	22.9\%.}
	\label{fig:q148}
\end{figure}
\begin{figure}[t]
	\script{q148_q1248_compare.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/q148_q1248_compare.pdf}
	\caption{Distributions of $\log_{10}$ difference in mismatch. The
	distribution labeled \texttt{q148} uses training waveforms listed in
	Tab.~\ref{tab:q148} while the \texttt{q1248} distribution uses waveforms
	listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Mismatches are calculated
	using the constant PSD with the loss function $\mathcal{L}_{\mathrm{mean}}$.}
	\label{fig:q148_q1248_compare}
\end{figure}

@@

\begin{figure}[t]
	\script{all_quadrants.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
	\caption{Distributions of mismatches in the top-left and bottom-right regions.
	We use a constant noise spectrum to calculate mismatch and
	$\mathcal{L}_{\mathrm{mean}}$ for the loss function. We see waveforms in the top-left region generally improves while waveforms in the bottom-right region worsened. \kl{I tried to overlap these two histograms together but it makes them very hard to read. }}
	\label{fig:all_quadrants}
\end{figure}

\begin{figure}[t]
	\script{ps_q148_quadrants.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
	\caption{\kw{The color scheme makes the plot a bit hard to read.}Parameter
	space of testing waveforms. Each region is fitted independently. Colorbar
	represents log difference of mismatches before and after optimization.}
	\label{fig:ps_q148_quadrant}
\end{figure}

>>>>

\begin{abstract}
    We present a simple and general method of recalibrating gravitational wave
    (GW) phenomenological waveform models jointly. By using {\jax} and
    {\ripple}, we can perform automatic differentiation to functions, which
    allows us to use gradient-based optimization methods to recalibrate waveform
    coefficients in IMRPhenomD model. This method reduces systematic bias
    previously introduced to the model and generally can improve waveform
    accuracy. With recalibrated coefficients, we found that the typical
    \textit{mismatch} has a $50\%$ decrease. Furthermore, we analyze the
    accuracy base on the waveform's intrinsic parameters. We found that waveform
    accuracy has significant dependence on black hole spin. Reduced spin
    approximation introduces degeneracy in spin, which prevented further
    improvement. We isolated regions in the parameter space that does not fit
    the waveform ansatz. These results allow us to understand more about how to
    develop newer phenomenological models. 
\end{abstract}

\section{Introduction} \label{sec:intro}

@@

\begin{figure}[t]
	\script{q148.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/q148.pdf}
	\caption{Distributions of waveform mismatches calculated using both
	$\mathcal{L}_{\mathrm{mean}}$ and $\mathcal{L}_{\mathrm{fl}}$. We use 
	training waveforms listed in Tab.~\ref{tab:q148} and
	mismatches are weighted with the constant PSD. For the
	$\mathcal{L}_{\mathrm{mean}}$ distributions, the median (dashed lines) decreased by 50.0\%
	while the median of $\mathcal{L}_{\mathrm{fl}}$ distribution decreased by
	22.9\%.}
	\label{fig:q148}
\end{figure}
\begin{figure}[t]
	\script{q148_q1248_compare.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/q148_q1248_compare.pdf}
	\caption{Distributions of $\log_{10}$ difference in mismatch. The
	distribution labeled \texttt{q148} uses training waveforms listed in
	Tab.~\ref{tab:q148} while the \texttt{q1248} distribution uses waveforms
	listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Mismatches are calculated
	using the constant PSD with the loss function $\mathcal{L}_{\mathrm{mean}}$.}
	\label{fig:q148_q1248_compare}
\end{figure}

@@

\begin{figure}[t]
	\script{all_quadrants.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
	\caption{Distributions of mismatches in the top-left and bottom-right regions.
	We use a constant noise spectrum to calculate mismatch and
	$\mathcal{L}_{\mathrm{mean}}$ for the loss function. Waveforms in the top-left region generally improves while waveforms in the bottom-right region worsened, as indicated by the median (dashed lines). \kl{I tried to overlap these two histograms together but it makes them very hard to read. }}
	\label{fig:all_quadrants}
\end{figure}

\begin{figure}[t]
	\script{ps_q148_quadrants.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
	\caption{Parameter space of testing waveforms. Each region is fitted independently. Colorbar
	represents log difference of mismatches before and after optimization.}
	\label{fig:ps_q148_quadrant}
\end{figure}
